---
title: "tidygate: high-level data analysis and manipulation in tidyverse style."
output: github_document
---
**It tidies up your playground!**


<!---

[![Build Status](https://travis-ci.org/stemangiola/tidygate.svg?branch=master)](https://travis-ci.org/stemangiola/tidygate) [![Coverage Status](https://coveralls.io/repos/github/stemangiola/tidygate/badge.svg?branch=master)](https://coveralls.io/github/stemangiola/tidygate?branch=master)

-->

<!-- badges: start -->
  [![Lifecycle:maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
 <!-- badges: end -->





## Minimal input data frame

element | feature | value
------------ | ------------- | -------------
`chr` or `fctr` | `chr` or `fctr` | `numeric`

## Output data frame

element | feature | value | gate
------------ | ------------- | ------------- | -------------
`chr` or `fctr` | `chr` or `fctr` | `numeric` | ...


```{r, echo=FALSE, include=FALSE, }
library(knitr)
knitr::opts_chunk$set(cache = TRUE, warning = FALSE,
                      message = FALSE, cache.lazy = FALSE)

library(tidyverse)
library(magrittr)
library(tidygate)

```

## Installation

```{r, eval=FALSE}

devtools::install_github("stemangiola/tidygate")

```

## Introduction

tidygate is a collection of wrapper functions for high level data analysis and manipulation following the tidy paradigm.

## Tidy data

```{r}
mtcars_tidy = 
	mtcars %>% 
	as_tibble(rownames="car_model") %>% 
	mutate_at(vars(-car_model,- hp, -vs), scale) %>%
	gather(feature, value, -car_model, -hp, -vs)

mtcars_tidy
```

## `ADD` versus `GET` versus `ONLY` modes

Every function takes a tidyfeatureomics structured data as input, and (i) with action="add" outputs the new information joint to the original input data frame (default), (ii) with action="get" the new information with the element or feature relative informatin depending on what the analysis is about, or (iii) with action="only" just the new information. For example, from this data set

```{r, cache=TRUE}
  mtcars_tidy
```

**action="add"** (Default)
We can add the MDS dimensions to the original data set

```{r, cache=TRUE, eval=FALSE}
  mtcars_tidy %>%
    reduce_dimensions(
    	car_model, feature, value, 
    	method="MDS" ,
    	.dims = 3,
    	action="add"
    )
```

**action="get"** 
We can add the MDS dimensions to the original data set selecting just the element-wise column

```{r, cache=TRUE, eval=FALSE}
  mtcars_tidy %>%
    reduce_dimensions(
    	car_model, feature, value, 
    	method="MDS" ,
    	.dims = 3,
    	action="get"
    )
```

**action="only"**
We can get just the MDS dimensions relative to each element

```{r, cache=TRUE, eval=FALSE}
  mtcars_tidy %>%
    reduce_dimensions(
    	car_model, feature, value, 
    	method="MDS" ,
    	.dims = 3,
    	action="only"
    )
```